---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# valid

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/valid)](https://CRAN.R-project.org/package=valid)
<!-- badges: end -->

## Installation

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("rappster/valid")
```

## What?

Helpers for facilitating systematic input validation

## Why?

Input validation is tedious already. This package hopefully provides some
functionality to make the task a bit easier or at least more fun.

## How?

```{r}
library(valid)
```

The actual workhorse is `valid()` while `valid2()` makes choice input a bit more
convenient via the `...` mechanism.

### Built-in functions

There are some built-in functions for typical validation tasks.

The number of such built-in functions will grow over time in a "scratch my own itch" manner

```{r}
valid_authentication()
valid_authentication("https")
valid_authentication(2)
try(valid_authentication("I don't exist"))
try(valid_authentication(3))

valid_dep_types()
valid_dep_types("Suggests")
try(valid_dep_types("I don't exist"))

valid_devops_envs()
valid_devops_envs("staging")
try(valid_devops_envs("I don't exist"))

valid_licenses()
valid_licenses("mit")
try(valid_licenses("I don't exist"))
```

You can reverse the order or flip names and values

```{r}
valid_devops_envs(reverse = TRUE)
valid_devops_envs(flip = TRUE) # doesn't really make sense in this case as both is lowercase
valid_devops_envs(unname = TRUE)
```

### Custom function

#### Definition

You can build your functions on top of `valid::valid()` 

```{r}
my_valid_devops_envs <- function(
  x = character(),
  ...
) {
  choices <- c("dev", "staging", "prod")
  names(choices) <- toupper(choices)

  valid(
    choice = x,
    choices = choices,
    ...
  )
}

my_valid_devops_envs2 <- function(
  ...
) {
  choices <- c("dev", "staging", "prod")
  names(choices) <- toupper(choices)

  valid2(
    ...,
    .choices = choices
  )
}
```

#### Apply without explicit choice

```{r}
my_valid_devops_envs()
```

```{r}
my_valid_devops_envs(reverse = TRUE)
```
```{r}
my_valid_devops_envs(flip = TRUE)
```
```{r}
my_valid_devops_envs(unname = TRUE)
```
```{r}
my_valid_devops_envs(flip = TRUE, unname = TRUE)
```

#### Choice via name

Valid:

```{r}
my_valid_devops_envs("PROD")
```

```{r}
my_valid_devops_envs(c("STAGING", "PROD"))
my_valid_devops_envs2("STAGING", "PROD")
```

Invalid:

```{r}
try(my_valid_devops_envs("ABC"))
```

```{r}
my_valid_devops_envs("ABC", strict = FALSE)
```

Partially invalid: 

```{r}
try(my_valid_devops_envs(x = c("DEV", "ABC")))
try(my_valid_devops_envs2("DEV", "ABC"))
```

```{r}
my_valid_devops_envs(x = c("DEV", "ABC"), strict = FALSE)
my_valid_devops_envs2("DEV", "ABC", .strict = FALSE)
```

#### Choice via value

Valid:

```{r}
my_valid_devops_envs("dev")
```

```{r}
my_valid_devops_envs(c("dev", "staging"))
my_valid_devops_envs2("dev", "staging")
```

Invalid:

```{r}
try(my_valid_devops_envs("abc"))
```

```{r}
my_valid_devops_envs("abc", strict = FALSE)
```

Partially invalid

```{r}
try(my_valid_devops_envs(c("dev", "abc")))
try(my_valid_devops_envs2("dev", "abc"))
```

```{r}
my_valid_devops_envs(c("dev", "abc"), strict = FALSE)
my_valid_devops_envs2("dev", "abc", .strict = FALSE)
```

#### Choice via index/position

Valid:

```{r}
my_valid_devops_envs(2)
```

```{r}
my_valid_devops_envs(c(2, 3))
my_valid_devops_envs2(2, 3)
```

Invalid:

```{r}
try(my_valid_devops_envs(4))
```

```{r}
my_valid_devops_envs(4, strict = FALSE)
```

Partially invalid: 

```{r}
try(my_valid_devops_envs(x = c(1, 4)))
try(my_valid_devops_envs2(1, 4))
```

```{r}
my_valid_devops_envs(x = c(1, 4), strict = FALSE)
my_valid_devops_envs2(1, 4, .strict = FALSE)
```

### Use validation function inside of another function

```{r}
foo <- function(
  devops_env = my_valid_devops_envs2("dev")
) {
  # Input validation
  devops_env <- match.arg(devops_env, my_valid_devops_envs2())
  
  # Body
  stringr::str_glue("DevOps env: {stringr::str_c(devops_env, collapse = ', ')}")
}
```

#### Keeping the default

```{r}
foo()
```

#### Eager input validation

```{r}
foo(
  devops_env = my_valid_devops_envs2("dev")
)
```

```{r}
try(
  foo(
    devops_env = my_valid_devops_envs2("abc")
  )
)
```

```{r}
try(
  foo(
    devops_env = my_valid_devops_envs2("dev", "abc")
  )
)
```

#### Lazy input validation

```{r}
foo(
  devops_env = "dev"
)
```

```{r}
try(
  foo(
    devops_env = "abc"
  )
)
```

## Code of Conduct

Please note that the valid project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

