---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# valid

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/valid)](https://CRAN.R-project.org/package=valid)
<!-- badges: end -->

## Installation

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("rappster/valid")
```

## What?

Helpers for facilitating systematic input validation

## Why?

Input validation is tedious already. This package hopefully provides some
functionality to make the task a bit easier or at least more fun.

## Overview

```{r}
library(valid)
```

### Some build-in functions

The amount of such built-in functions will grow over time in a "scratch my own itch" manner

```{r}
valid_dep_types()
valid_dep_types("Suggests")
try(valid_dep_types("I don't exist"))

valid_authentication()
valid_authentication("https")

valid_licenses()
valid_licenses(flip = TRUE)
```

### Custom function

#### Definition

You can build your functions on top of `valid::valid()` 

```{r}
#' Valid devops environments
#'
#' @param x [[character]] DevOps environment
#' @param reverse [[logical]] Reverse order
#' @param flip [[logical]] Flip values and names
#' @param strict [[logical]] Throw error if invalid choice or return empty
#'
#' @return
#' @export
#'
#' @examples
valid_devops_envs <- function(
  x = character(),
  reverse = FALSE,
  flip = FALSE,
  strict = TRUE
) {
  values <- c("dev", "prod")
  names(values) <- toupper(values)

  valid(
    choice = x,
    choices = values,
    reverse = reverse,
    flip = flip,
    strict = strict
  )
}
```

#### Apply without explicit choice

```{r}
valid_devops_envs()
```

```{r}
valid_devops_envs(reverse = TRUE)
```
```{r}
valid_devops_envs(flip = TRUE)
```

#### Choice via value

Valid:

```{r}
valid_devops_envs("dev")
```
Invalid:

```{r}
try(valid_devops_envs("abc"))
```

Invalid but non-strict:

```{r}
valid_devops_envs("abc", strict = FALSE)
```

#### Choice via name

Valid:

```{r}
valid_devops_envs("PROD")
```

Invalid:

```{r}
try(valid_devops_envs("ABC"))
```

Invalid but non-strict:

```{r}
valid_devops_envs("ABC", strict = FALSE)
```

#### Choice via index/position

Valid:

```{r}
valid_devops_envs(2)
```

Invalid:

```{r}
try(valid_devops_envs(3))
```

Invalid but non-strict:

```{r}
valid_devops_envs(3, strict = FALSE)
```
### Use validation function inside of another function

```{r}
foo <- function(
  devops_env = valid_devops_envs("dev")
) {
  # Input validation
  devops_env <- match.arg(devops_env, valid_devops_envs())
  
  # Body
  stringr::str_glue("DevOps env: {stringr::str_c(devops_env, collapse = ', ')}")
}
```

#### Keeping the default

```{r}
foo()
```

#### Eager input validation

```{r}
foo(
  devops_env = valid_devops_envs("dev")
)
```

```{r}
try(
  foo(
    devops_env = valid_devops_envs("abc")
  )
)
```

#### Lazy input validation

```{r}
foo(
  devops_env = "dev"
)
```

```{r}
try(
  foo(
    devops_env = "abc"
  )
)
```

## Code of Conduct

Please note that the valid project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

